---
title: "Reproducible Document"
author: "Emma"
format: html
editor_options: 
  chunk_output_type: inline
---

## Setup

```{r setup}
#| message: false
#| warning: false

library(tidyverse)

post_meal_data <- read.csv2(here::here("data/post-meal-insulin.csv"))

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)

tidier_post_meal_data <- small_post_meal_data %>%
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  ) %>%
  filter(glykemi == 0) %>%
  select(-glykemi)

tidied_post_meal_data <- tidier_post_meal_data %>%
  mutate(
    insulin_glucose_ratio = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First-degree relative")
  )
```


## Writing R code 

```{r}
2 + 2
```

## About me
- Emma
- SDU

My name is **Emma** and i live in *odense*. 

## Simple code
```{r}
3 * 3
```

## Testing for git

The weather is shit

## Showing the data
```{r}
post_meal_data

glimpse(post_meal_data)
```

## Plot one continues variable 
```{r}
#| fig-cap: "Distribution of BMI."
#| label: fig-bmi-histo

ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

In figure @fig-bmi-histo you can see the BMI in a histogram

## Exercise: Discrete plots 

```{r}
#| fig-cap: "Barplot of Group"
#| label: fig-group-bar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```
See @fig-group-bar above for a cool plot!

## plotting two discrete variables

```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

## Side by side plots

```{r fig-bmi-glycemia}
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#| - "Distribution of BMI"
#| - "Number of those with glycemia"
#| layout-ncol: 2

ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting two continuous variables

```{r}
#| fig-cap: "Scatterplot of BMI and AUC for glucose"
#| label: fig-bmi-auc-pg

ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

## Selecting data
```{r}
select(post_meal_data, Age)
select(post_meal_data, Age, BMI, Weight)
select(post_meal_data, -Group)
select(post_meal_data, starts_with("PG"))
select(post_meal_data, ends_with("DXA"))
select(post_meal_data, contains("0"))
```

## Renaming Colums

```{r}
small_post_meal_data %>%
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```

## Select and rename columns 

```{r}
# Task 1.
post_meal_data %>% select(contains("OGTT"))

# Task 2.
post_meal_data %>% select(contains(".."), -contains("..."))

# Task 3.
post_meal_data %>% rename(P.Glucose.Minus.5.OGTT = P.Glucose..5.OGTT, Insulin.Minus.5.OGTT.X = Insulin..5.OGTT.X)

# Task 4.
post_meal_data %>%
  select(id, BMI, Length, Weight, Age) %>%
  rename(Height = Length)
```

## Filtering data

```{r}
small_post_meal_data %>%
  filter(glykemi == 1)

small_post_meal_data %>%
  filter(BMI >= 25)

small_post_meal_data %>%
  filter(Group == "CTR")

small_post_meal_data %>%
  filter(Group == "CTR" | BMI >= 25)
```

## Modifying Data

```{r}
tidier_post_meal_data %>%
  mutate(insulin_glucose_ratio = auc_insulin / auc_glucose)

tidier_post_meal_data %>%
  mutate(
    insulin_glucose_ratio = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First-degree relative")
  )
```
